Merging datasets through deep learning

### 2018-08-16 paper2018/introduction.tex
adding another section

Merging two datasets using joins is a key operation in data management.  Data management systems have largely focussed solely on equi-joins, where equality of strings or numeric values of the two columns determines if rows should be joined.  However, equality at a string level is inadequate because the same entity may in many cases be expressed with slight variations of the same name, as shown in Table~\ref{table-example} (e.g., \textit{Douglas Adams} and \textit{Douglas Noel Adams}).  This problem occurs for many entity types such as company names, addresses, product descriptions, and conference venues.

A common technique to overcome this problem is to perform fuzzy joins using string similarity algorithms such as edit-distance, Jaro-Winkler and TF-IDF (e.g., \cite{Cohen2003}).  In applying fuzzy joins, a key problem is to avoid quadratic comparisons between pairs of strings drawn from the two columns.  A blocking strategy is used to reduce the number of pairs for fuzzy joins, such that only strings that share a common prefix, or suffix will be compared further.  String matching algorithms often do not work because transformations of entity names in perfectly valid ways can yield very different strings (e.g., switching the last name and first name of a person, or providing acronyms for a company name).

@@

In this paper, we propose a novel data driven approach to the problem of joining different representations of the same entity.  Our approach relies on building supervised deep learning models to automatically learn the correct set of transformations needed to compute the equality of two cell values.  For the closely related problem of face recognition and person re-identification, deep learning networks have been used successfully to discriminate pairs that represent the same entity from those that don't.  However, we know of no work that has adapted these techniques to the problem of fuzzy joins for entity names.

Because a deep learning model can learn the features necessary for matching from the data, it should in theory be able to generalize better to new datasets than the data driven approaches outlined in \cite{He:2015:SJS:2824032.2824036}.   Conceptually, our approach is similar in spirit to the approach described in \cite{auto-join-joining-tables-leveraging-transformations}, but we use deep neural networks to learn the right function for the join operation instead of synthesizing different programs for different entity types.  The advantage of building such a function over the approach described in \cite{auto-join-joining-tables-leveraging-transformations} is that we can use this function in novel ways to completely eliminate the blocking step that is typically needed to identify the set of potentially joinable rows, as we describe below.

Our solution to the join problem involves building a system that performs deep metric learning.  That is, the specific function the network learns is to map input vectors corresponding to same entity closer in vector space compared to vectors that correspond to different entities.  As in face recognition, we do this by training a `triplet siamese network'.  However, entity names have very different properties from faces, so we needed to adapt the approaches used in face recognition to the entity matching problem with a novel method to choose the sample for training, and novel loss functions as objectives for the metric learning problem.

At training, the network is given a triplet consisting of an `anchor' (e.g. \textit{Douglas Adams}), a positive element (e.g. \textit{Douglas Noel Adams}) and a negative element (e.g., John Adams), and it learns to produce a small distance estimate for the positive pair (close to 0), and a large distance estimate for the negative pair, using an objective function. A classification form of this function that simply learns whether two names belong to the same entity can be used to perform fuzzy joins, but then we would need a blocking strategy analogous to what has been used in the literature to identify the set of joinable pairs \cite{auto-join-joining-tables-leveraging-transformations}.

Our observation is that one can actually exploit what the siamese network produces to eliminate this blocking step altogether.  Specifically, the last hidden layer of the siamese network is effectively a `vector embedding' that contains the critical features needed for computing the distance estimate.  If such models are pre-built for a given entity type, we can then use the models to produce vector embeddings for all cell values in the two columns to be joined in a single linear pass, and index just the second column values in an approximate nearest neighbors index.  Then, for each cell value in the first column to be joined, querying the k-nearest neighbors should provide the set of values in the second column that can be joined.  Because approximate nearest neighbor algorithms have been applied successfully to billions of items \cite{JDH17}, this approach should scale just as well as join algorithms based on blocking, but has the added advantage that the blocking step is also data driven, and hence can adapt to different entity types very effectively.

\begin{table}[!htb]
    \caption{Example of a join problem}
    \begin{subtable}{.5\linewidth}
      \centering
        \caption{Table A}
        \begin{tabular}{|l|l|}
          \hline
           Dept & Name \\
           \hline
           1    & Douglas Adams \\
           2    & John Adams \\
           3  & Douglas Bard \\
           \hline
        \end{tabular}
    \end{subtable}%
    \begin{subtable}{.5\linewidth}
      \centering
        \caption{Table B}
        \begin{tabular}{|l|l|}
          \hline
           Name & Salary \\
           \hline
           Douglas N. Adams & 2000 \\
           Adams, John & 3000 \\
           John Smith & 3000 \\
           \hline
        \end{tabular}
    \end{subtable}
    \label{table-example}
\end{table}

Our key contributions are as follows:
\begin{itemize}
\item We demonstrate that triplet loss based siamese networks can be used to learn a distance function that successfully discriminates same entity pairs from different pairs with an accuracy of 98\% and 97\% for about 200,000 people's names and 70,000 company names respectively, drawn from Wikidata.  More importantly, we show that such a network has a high recall of 83\% and 79\%, finding the variants of an entity name in the top 20 neighbors for people and companies respectively.  Moreover, within the 20 neighbors, 77\% and 75\% of the positive names were closer to the anchor than negative names for people and companies respectively.  Our results suggest that deep learning models can be used successfully for data driven joins, assuming that such models are built for common entity types.
\item Training neural networks for deep metric learning is difficult because one cannot show quadratic combinations of the data to the system.  In face recognition, a number of techniques exist for triplet selection, and methods for selecting triples is an active area of research.  We developed a new technique for triplet selection for entity names because the techniques from face recognition do not generalize well to entity names.
\item We also developed a novel loss function to train the network to discriminate entity names from names of other entities.  We compare the results for this loss function with the ones that have been used for face recognition, and show that this loss function can perform much better than the existing functions for deep learning in this domain.
\end{itemize}

The rest of the paper is organized as follows: Section~\ref{siamese networks} describes how we adapted siamese networks to the problem of entity matching, specifically the problem of triplet selection for entity names, and the need for a new type of loss function to improve metric learning for this problem, section ~\ref{dataset} describes characteristics of the datasets we used for training, as well as cleansing and data augmentation techniques we used to generate data, section ~\ref{results} provides the experimental results, section~\ref{related work} covers related work, and section~\ref{conclusions} describes future work and conclusions.

>>>>

Merging datasets is a key operation for data analytics. A frequent requirement for merging datasets is a join between two columns that frequently have different surface representations for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}, as shown in Table~\ref{table-example}.  This problem occurs for many entity types such as people's names, company names, addresses, product descriptions, conference venues, or even people's faces.  Data management systems have however, largely focussed solely on equi-joins, where string or numeric equality determines which rows should be joined, because such joins are efficient.

Extensions in data management systems for handling `fuzzy' joins across different surface forms for the same entity typically use string similarity algorithms such as edit-distance, Jaro-Winkler and TF-IDF (e.g., \cite{Cohen2003}).  In applying fuzzy joins, a key problem is to avoid quadratic comparisons between pairs of strings drawn from the two columns.  A blocking strategy is used to reduce the number of pairs for fuzzy joins, such that only strings that share a common prefix, or suffix will be compared further.  String matching algorithms often do not work however, because transformations of entity names in perfectly valid ways can yield very different strings.  As shown in Table~\ref{table-example}, \textit{John Smith} is more similar by string similarity to \textit{John Adams} than \textit{Adams, John}, but of course \textit{Adams, John} is the valid alternate surface form for \textit{John Adams}.

@@

In this paper, we propose a novel data driven approach to the problem of joining different surface representations of the same entity.  Our approach relies on building supervised deep learning models to automatically learn the correct set of transformations needed to compute the equality of two cell values.  Because a deep learning model can learn the features necessary for matching from the data, it should in theory be able to generalize better to new datasets than the data driven approaches outlined in \cite{He:2015:SJS:2824032.2824036}.   Conceptually, our approach is similar in spirit to the approach described in \cite{auto-join-joining-tables-leveraging-transformations}, but we use deep neural networks to learn the right function for the join operation instead of synthesizing different programs for different entity types.  The advantage of building such a function compared to work described in \cite{auto-join-joining-tables-leveraging-transformations} is that (a) we can use this function in novel ways to completely eliminate the blocking step that is typically needed to identify the set of potentially joinable rows, as we describe below, and (b) we can use this function on arbitrary surface forms such as faces, where valid transformations of a face cannot be readily described in terms of a program.

Specifically, our solution to the join problem involves building a system that performs deep metric learning.  That is, the network learns is to map input vectors corresponding to same entity closer in vector space compared to vectors that correspond to different entities.  For the closely related problem of face recognition and person re-identification, deep learning networks have been used successfully to discriminate pairs that represent the same entity from pairs that do not.  However, simply building a model that discriminates pairs of entities is not sufficient; if used naively, this would mean a quadratic comparison between cell values in the two columns to be joined.  We know of no work that has adapted these techniques to the problem of how to use such deep learning models to perform joins efficiently for different surface forms of the same entity.

As in face recognition, our system for metric learning is built by training a `triplet siamese network' to learn to map different surface forms of the same entity closer in vector space than different forms of the same entity.  Entity names have very different properties from faces, so we needed to adapt the approaches used in face recognition to the problem of matching entity names with a novel method to choose the sample for training, and novel loss functions as objectives for the metric learning problem.  At training, the network is given a triplet consisting of an `anchor' (e.g. \textit{Douglas Adams}), a positive element (e.g. \textit{Douglas Noel Adams}) and a negative element (e.g., John Adams), and it learns to produce a small distance estimate for the positive pair (close to 0), and a large distance estimate for the negative pair, using an objective function. A classification form of this function that simply learns whether two names belong to the same entity can be used to perform fuzzy joins, but then we would need a blocking strategy analogous to what has been used in the literature to identify the set of joinable pairs \cite{auto-join-joining-tables-leveraging-transformations}.

Our observation is that one can actually exploit what the siamese network produces to eliminate this blocking step altogether.  Specifically, the last hidden layer of the siamese network is effectively a `vector embedding' that contains the critical features needed for computing the distance estimate.  If such models are pre-built for a given entity type, we can then use the models to produce vector embeddings for all cell values in the two columns to be joined in a single linear pass, and index just the second column values in an approximate nearest neighbors index.  Then, for each cell value in the first column to be joined, querying the k-nearest neighbors should provide the set of values in the second column that can be joined.  Because approximate nearest neighbor algorithms have been applied successfully to billions of items \cite{JDH17}, this approach should scale just as well as join algorithms based on blocking, but it has the added advantage that the blocking step is also data driven, and hence can adapt to different entity types very effectively.

\begin{table}[!htb]
    \caption{Example of a merge problem}
    \begin{subtable}{.5\linewidth}
      \centering
        \caption{Table A}
        \begin{tabular}{|l|l|}
          \hline
           Dept & Name \\
           \hline
           1    & Douglas Adams \\
           2    & John Adams \\
           3  & Douglas Bard \\
           \hline
        \end{tabular}
    \end{subtable}%
    \begin{subtable}{.5\linewidth}
      \centering
        \caption{Table B}
        \begin{tabular}{|l|l|}
          \hline
           Name & Salary \\
           \hline
           Douglas N. Adams & 2000 \\
           Adams, John & 3000 \\
           John Smith & 3000 \\
           \hline
        \end{tabular}
    \end{subtable}
    \label{table-example}
\end{table}

Our key contributions are as follows:
\begin{itemize}
\item We demonstrate that triplet loss based siamese networks can be used to learn a distance function that successfully discriminates same entity pairs from different pairs with an accuracy of 98\% and 97\% for about 200,000 people's names and 70,000 company names respectively, drawn from Wikidata.  More importantly, we show that such a network has a high recall of 83\% and 79\%, finding the variants of an entity name in the top 20 neighbors for people and companies respectively.  Moreover, within the 20 neighbors, 77\% and 75\% of the positive names were closer to the anchor than negative names for people and companies respectively.  Our results suggest that deep learning models can be used successfully for data driven joins, assuming that such models are built for common entity types.
\item Training neural networks for deep metric learning is difficult because one cannot show quadratic combinations of the data to the system.  In face recognition, a number of techniques exist for triplet selection, and methods for selecting triples is an active area of research.  We developed a new technique for triplet selection for entity names because the techniques from face recognition do not generalize well to entity names.
\item We also developed a novel loss function to train the network to discriminate entity names from names of other entities.  We compare the results for this loss function with the ones that have been used for face recognition, and show that this loss function can perform much better than the existing functions for deep learning in this domain.
\item We demonstrate how such models can then be used to perform joins across different surface forms of the same entity efficiently, with a complexity of $\mathcal{O}(n\log{}n)$.  
\end{itemize}

The rest of the paper is organized as follows: Section~\ref{siamese networks} describes how we adapted siamese networks to the problem of entity matching, specifically the problem of triplet selection for entity names, and the need for a new type of loss function to improve metric learning for this problem, section ~\ref{joins} describes how to use these models for efficient joins across datasets, section ~\ref{dataset} describes characteristics of the datasets we used for training, as well as cleansing and data augmentation techniques we used to generate data, section ~\ref{results} provides the experimental results, section~\ref{related work} covers related work, and section~\ref{conclusions} describes future work and conclusions.

### 2018-08-16 paper2018/main.tex
changed notation for schroff loss
adding another section

\def\year{2018}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai18}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\usepackage{subcaption}

\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{
/Title (Fuzzy joins through deep learning)
/Author (Kavitha Srinivas, Yehuda Gale, Julian Dolby)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Fuzzy joins through deep learning}

@@

\maketitle
\begin{abstract}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions. 
\end{abstract}

\input{introduction}
\input{siamese_networks}

>>>>

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\usepackage{subcaption}

\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{
/Title (Merging datasets through deep learning)
/Author (Kavitha Srinivas, Yehuda Gale, Julian Dolby)}
\setcounter{secnumdepth}{0}  
\usepackage{bm}

 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Merging datasets through deep learning}

@@

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent requirement for merging datasets is a join between two columns that frequently have different surface forms for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}).  Data management systems are currently limited to performing merges based on string equality, or at best use string similarity to allow so-called `fuzzy' joins.  We explore a novel approach to performing merges based on recent advances in deep learning.  Our approach builds a deep learning model that maps names into a set of vector embeddings such that names for the same entity are closer in vector space than names for a different entity.  We then index these vector embeddings in an approximate nearest neighbors algorithm to find the names that can be potentially joined together.  Our approach for merging datasets is efficient $\mathcal{O}(n\log{}n)$ and flexible because it can be applied to any data that can be mapped into vector space with the constraint that joinable surface forms are closer in vector space than disparate forms.  We demonstrate the feasibility of the approach for people's names and company names drawn from Wikidata.
\end{abstract}

\input{introduction}
\input{siamese_networks}
\input{joins}
